---
title: "Homework_5"
author: "Andrea_Sanchez_Hidalgo"
date: "2025-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, Part 1 README, eval=TRUE}
# Create README.md file from R

readme_text <- "

# Homework 5

This README file contain instructions for HW5. 

## Description
The purpose of this HW is learn to commmit to GIthub in R.

## Requirements or Packages

## Instructions
Explain how to run the analysis or scripts.

"

# Write the file to the project root
writeLines(readme_text, 'README.md')
```


```{r, Recreating a graph}
##It shows monthly homicides in Baltimore, with a reference added for the date of the arrest of Freddie Gray and color used to show colder months (November through April) versus warmer months (May through October). There is a smooth line added to help show seasonal and long-term trends in this data.

##Library load
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(lubridate)

##read file
getwd()
homicides <- read.csv ("https://raw.githubusercontent.com/AndreaSH15/EHRS-ASH/refs/heads/main/homicide-data.csv")

View(homicides)
str(homicides)

##extract infromation from Baltimore
homicides_baltimore <- homicides %>%
  filter(city == "Baltimore") %>%
  mutate(
    reported_date = ymd(reported_date),
    year = year(reported_date),
    month = month(reported_date),
    season = if_else(month %in% c(11,12,1,2,3,4), "colder", "warmer")
  ) %>%
  group_by(year, month, season, disposition) %>%
  summarise(homicides = n(), .groups = "drop")

homicides_baltimore_monthly <- homicides_baltimore %>%
  group_by(year, month, season) %>%
  summarise(total_homicides = sum(homicides), .groups = "drop")

View(homicides_baltimore)
```

##Freddie Gray

Interestingly, this database does not contain the name of Freddy Gray as a victim, however, according to Wikipedia, he was arrested on April 12, 2015, for possession of a knife, however, he was attacked by police officers on the same date and died in R Adams Cowley Shock Trauma Center. Gray's death on April 19, 2015, was ascribed to injuries to his cervical spinal cord. 

Source: https://en.wikipedia.org/wiki/Killing_of_Freddie_Gray

Note: As he was not in the dataset obtained, the date of the arrest used on this excercise (April 12, 2015) was taken from Wikipedia.

```{r, replication of the graph}

```

